---
title: "Setup Project"
author: "Sebastian DiGeronimo"
date: "2022-11-22"
output: html_document
---

# Step 1: Load/install necessary packages for setting up .Rprofile file
```{r setup, include=FALSE}
if (!require("librarian")) {
    install.packages("librarian")
}

librarian::shelf(
    librarian, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional
    cli, rlang, data.table, geosphere, ggforce, hablar, hms, httr, janitor, jsonlite, 
    knitr, lubridate, iobis/obistools, plotly, readxl, parzer,
    renv, rmarkdown, tools, vroom, worrms
)

if (!require("conflicted")) {
    install.packages("conflicted")
}

library("conflicted")

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

# Step 2: Run `rprofile_setup()`
Creates .Rprofile file that contains:
1. Designation of cloud storage path 
2. Loading of libraries needed for use in most processes
3. Attaches functions and scripts to search_path

NOTE: to take effect, will need to restart R afterwards
```{r}
source(here("scripts", "rprofile_setup.R"))
prof <- here()

# .choose 
#   TRUE: gives popup to select directory
#   FALSE: open .Rprofile for manual addition to line 1
rprofile_setup(prof = prof, 
               .choose = TRUE)

rm(list = ls())
```

# Step 3: Use .Rprofile to setup directories, source scripts with functions needed and attach to search path

1. Creates directory for consistent file structure

2. Copy files from designated location
    - The first time using, this will open a `.Rprofile` file and add `cloud_dir` if not already added. This is where the raw data would be stored. If you want to select through a popup set `copy_files_cloud(.choose = TRUE)` and will paste to .Rprofile. You can manually paste into .Rprofile if `.choose = FALSE`. You will then need to restart R.
    - If you want to auto-look for new files, edit `copy_files_cloud(ask = TRUE)` to `copy_files_cloud(ask = TRUE, auto = TRUE)`.
    Each `Restart` in R will look for new files and ask if you want to download. 

3. Load functions for searching the raw for taxa names and connecting it to a master sheet
- if a master sheet doesn't exists locally, it will start a new one and put it in:
    - `data/metadata/aphia_id`

# Run Rmd file in order
1. `taxa_match_up_WoRMS.Rmd`
    - match raw data to taxa
2. `combine_zoo_logsheets.Rmd`
    - get metadata and merge all files together
3. `create_obis_event.Rmd`
    - combine raw data and metadata to create obis formatted data

