---
title: "Splitting Stations into Separate Files"
author: "Sebastian Di Geronimo"
date: "2024-02-16"
format: html
editor: source
---

# 1.0 ---- Summary of Document ----

This script uses the pre-OBIS zooplankton data and separates into identifier and
location ID. Then, the data will be save as a `.csv` with the intials of who 
identified the species and the location ID:
ex: "zooplankton_erddap_<NP|JRM>_<locationID>.csv"


This assumes that the data is stored locally and there is a variable called 
`cloud_dir` to show the path to Box to store the new files.

# 2.0 ---- Setup ----


## 2.1 Load Libraries

```{r setup, include=FALSE}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    quiet = TRUE,
    librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    
    # additional
    
)

conflicts_prefer(
  dplyr::filter(), 
  dplyr::select()
  )
  

```

## 2.2 Load Dataset

```{r load-data}
dat <-   
  here("data", "processed") %>%
  dir_ls(regexp = "zooplankton_erddap_20240209") %>%
  read_csv(show_col_types = FALSE) %T>% 
  print()

dat_nest <- nest(dat, .by = c("identifiedBy", "locationID")) %T>% print()

pwalk(
  list(
    dat_nest$identifiedBy,
    dat_nest$locationID,
    dat_nest$data,
    overwrite = FALSE # change to TRUE if want to overwrite data
  ),
  function(name, loc, data, overwrite = overwrite) {
    
      dat <-
      data %>%
      mutate(
        .before      = 1,
        identifiedBy = name,
        locationID   = loc
      )

    # "zooplankton_erddap_<NP|JRM>_<locationID>.csv"
    nam       <- if_else(str_detect(name, "Jaimie"), "JRM", "NLF")
    file_name <- glue("zooplankton_erddap_{nam}_{loc}.csv")
    dir_name  <- here(cloud_dir, "processed", "erddap")

    fs::dir_create(dir_name)

    if (!file_exists(here(dir_name, file_name))) {
      
      message(glue("Saving File {file_name}"))
      write_csv(
        x    = dat,
        file = here(dir_name, file_name),
        na   = ""
      )
      
    } else if (file_exists(here(dir_name, file_name)) && overwrite) {
      
      message(glue("Overwriting File {file_name}"))
      write_csv(
        x    = dat,
        file = here(dir_name, file_name),
        na   = ""
      )
      
    } else {
      
        message(glue("Skipping File {file_name}"))
    }
  }
)



```

