---
title: "Fix and Convert Previous Zooplankton Data"
author: "Tylar Murray"
date: "2023-04-14"
format: html
---

# 1.0 ---- Setup ----
## 1.1 Load Libraries
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    librarian, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here, conflicted,
    # broom # optional
    
    # additional
    cli, naniar, Hmisc
)

# shelf(conflicted) # may be needed if won't allow loading of certain packages

conflicts_prefer(
    dplyr::filter(), 
    dplyr::select()
    )
```

# 1.1 Make sure location of files is correct
From fix_prev_zoo_data.qmd
- create folders
  - place original data
  - place processed file to load here 
  
In case, you don't have these folder, creates them here
```{r create-folder}
base_dir <- here::here("data", "zoo_pre_2018")

fs::dir_create(here::here(base_dir, "processed"))
fs::dir_create(here::here(base_dir, "orig"))

```

## 1.2 Get File Paths
sp.csv from:
<https://drive.google.com/drive/folders/1IKBqpgOR6-bhJnm3wHwZAzvSYAA24Bsx>
put in "~/data/zoo_pre_2018"

```{r get-data-path}
# get file path to processed data
file_path <-
  here::here(base_dir, "processed") %>%
  fs::dir_ls(
    regexp  = "zoo_compiled_with_meta",
    recurse = TRUE
  ) %>%
  # function to check more recent made file to use
  last_mod() 

# read `.csv` file created by `fix_prev_zoo_data.Rmd`
dat <-
  file_path %>%
  # "../zoo_compiled_with_meta_20230411.csv" %>%
  
    readr::read_csv(
    file        = .,
    show_col_types = FALSE,
    guess_max   = 10000,
    name_repair = janitor::make_clean_names
  )

species_path <-
    here::here(base_dir) %>%
    fs::dir_ls(regexp = "sp\\.csv")

# read manually created verbatim name map `sp.csv`
species_map <-
  species_path %>%
  # "../sp.csv" %>%

  readr::read_csv(
    file = .,
    show_col_types = FALSE,
    name_repair = janitor::make_clean_names,
    na = c("", "NA", stringi::stri_dup("?", 1:5)) # ignore `?` 1 to 5 times
  ) %>%

  # select only necessary columns
  dplyr::select(
    jaimes_taxa_info,
    contains("sci_name"),
    lifestage
  )
```

# 2.0 ---- Merge Taxa Info ----
## 2.1 Merge Species Info with Raw Data
```{r load-csv-data}
# map Jaime's species names to WoRMS-readable using `sp.csv`
dat_merg <-
    species_map %>%

  # join on classification columns. `.` is where it gets piped into
  dplyr::left_join(
    mutate(dat, class2 = clasification, .after = clasification),
    .,
    by = join_by("clasification" == "jaimes_taxa_info")
  ) %>%
    
  # construct unique identifier for each row (used for occurrenceID later)
  dplyr::mutate(
    orig_data_row_ID = glue("{sheet_nm}:{clasification}:{lifestage}"),
    .before = 1
  ) %>%
  distinct()  %>%
    
  mutate(
    taxa_to_search =
      if_else(
        is.na(sci_name_guessed_by_tylar),
        clasification,
        sci_name_guessed_by_tylar
      )
  ) 
```


## 2.2 Check Duplicates for Unique IDs
```{r check-duplicates}
# validate: rows in file must be unique
# janitor::get_dupes(<dataframe name>, <column names to group together>) 

dupes <- janitor::get_dupes(dat_merg, orig_data_row_ID)

# Check if duplicated rows
if (nrow(dupes) > 0) {
    stop(
        glue("There are duplicated rows in merged data! ",
             "(n = {nrow(dupes)})")
        )
} else {
    message("No duplicated `orig_data_row_ID`.")
}

# only flowmeter data differs and one looks unrealistically large.

```

# 3.0 ---- Get AphiaID from WoRMS ----
Using `merge_taxa()` 
check = if want to check unmatched names
viewer = if want to see unmatched taxa using View()
```{r aphia-id}
# set location for aphiaID list and the base of the file name
file_exprs <- 
    file_expr(
    loc       = here::here(base_dir), 
    file_base = "aphia_taxa_jaime"
    )

# master_taxa_list(
#     taxa_list  = NULL, 
#     .cloud_dir = here(cloud_dir, "old_jaime_data"),
#     # file_base  = "aphia_taxa",
#     where_to   = "local", 
#     save       = FALSE,
#     .file_expr = file_exprs)

max_depth <-
  list(
    MR = 36.5,
    LK = 40.5,
    WS = 23
  )

taxa_base <-
  dat_merg %>%
  select(taxa_to_search) %>%
  distinct() %>%
  merge_taxa(.,
    .file_expr = file_exprs,
    check = FALSE,
    # check      = TRUE,
    use_cloud = FALSE,
    viewer = TRUE,
  )

taxa_full <-
  taxa_base %>%
  mutate(
    aphia_id = if_else(!is.na(scientificNameID),
      str_split(scientificNameID, ":",
        simplify = TRUE
      )[, 5],
      NA_character_
    ) %>%
      as.numeric(.),
    row = row_number(),
    info = pmap(
      list(
        aphia_id,
        taxa_orig,
        row
      ),
      # .progress = TRUE,
      .f =
        (\(.x, .y, .z)
        tryCatch(
          {
            if (.z == 1) {
              cat(sprintf(
                "\n%-4s | %-6s | names\n%s",
                "Row", "aphia",
                stringi::stri_dup("-", 37)
              ))
            }
            cat(sprintf("\n%-4d | %6s | %s | ", .z, .x, .y))
            x <- worrms::wm_record(.x)
            cat(x$scientificname)
            return(x)
          },
          error = function(e) {
            return(NULL)
          }
        ))
    )
  ) %>%
  unnest(info,
    names_repair = janitor::make_clean_names,
    keep_empty = TRUE
  ) %>%
  rename(
    "scientificName"   = valid_name,
    "scientificNameID" = lsid
  ) %>%
  mutate(
    scientificNameID = str_replace(
      scientificNameID, ":(\\d+)",
      glue(":{valid_aphia_id}")
    )
  ) %>%
  relocate(scientificName, scientificNameID, valid_aphia_id, .after = lifestage)


taxa_full 

# save updated version of base taxa and full taxa as .csv
if (FALSE) {
  write_excel_csv(
    x = taxa_base,
    file =  eval(file_exprs[[2]]),
    na = ""
  )
  write_excel_csv(
    x = taxa_full,
    file = here::here(
      base_dir,
      glue(
        "taxa_worms_info",
        format(Sys.Date(), "_%Y%m%d"),
        ".csv"
      )
    ),
    na = ""
  )
}

# update cloud service directory if have connection through explorer/Finder
if (FALSE && exists("cloud_dir") && !is.na(cloud_dir)) {
  master_taxa_list(
    taxa_list  = taxa_base,
    .cloud_dir = here(cloud_dir, "old_jaime_data"),
    where_to   = "cloud",
    save       = TRUE,
    .file_expr = file_exprs
  )
  master_taxa_list(
    taxa_list  = taxa_full,
    .cloud_dir = here(cloud_dir, "old_jaime_data"),
    where_to   = "cloud",
    save       = TRUE,
    .file_expr = "taxa_worms_info",
    sheet_name = "Full Taxonomic Info"
  )
}
```


```{r aphia-id-merge}
# filter out taxa names not needed
taxa_full_filt <- 
    taxa_full %>%
     filter(
         (str_detect(notes, "remove", negate = TRUE) | is.na(notes)) 
         & !is.na(scientificNameID))  
  

dat_merg_tx <- 
  # join on classification columns. `.` is where it gets piped into
    left_join(
      dat_merg,
      taxa_full_filt,
      by = c("taxa_to_search" = "taxa_orig")
  )  %>%
   mutate(
        .keep = "unused",
        .after = scientificName,
        lifestage = if_else(!is.na(lifestage.x), lifestage.x, lifestage.y)
        ) 

# round individual count. we don't know why there are fractional counts. 
dat_merg_tx <- 
    dat_merg_tx %>%
    
    mutate(
    
    ind_count = round(ind_count),
    # hardcode missing lat lon using averages in water-samples
    lon_in = case_when(
        is.na(lon_in) & str_detect(stn, "LK") ~ -81.41464542,
        is.na(lon_in) & str_detect(stn, "WS") ~ -81.71532752,
        is.na(lon_in) & str_detect(stn, "MR") ~ -80.37953386,
        TRUE ~ lon_in
    ),
    lat_in = case_when(
        is.na(lat_in) & str_detect(stn, "LK") ~ 24.53862793,
        is.na(lat_in) & str_detect(stn, "WS") ~ 24.47605506,
        is.na(lat_in) & str_detect(stn, "MR") ~ 25.00607727,
        TRUE ~ lat_in
    ),
    maximumDepthInMeters = case_when(
        str_detect(stn, "MR") ~ max_depth$MR,
        str_detect(stn, "LK") ~ max_depth$LK,
        str_detect(stn, "WS") ~ max_depth$WS
    )
) %>%
  
  # === set up to align with next chunk var names
  rename(
    
    "taxa_orig"        = sci_name_guessed_by_tylar,
    "individualCount"  = ind_count,
    "lifeStage"        = lifestage,
    "decimalLatitude"  = lat_in,
    "decimalLongitude" = lon_in,
  )

# eventID   <- "jaime_protocol_observation"
datasetID <- "USF_IMaRS_MBON_compiled_zoo_taxonomy_jaimie_2018"

```

# 4.0 ---- Construct Occurence ID ----
```{r construct-occurrence-table}
# map col names into dwc
occur <-
  dat_merg_tx %>%
  # left_join(event,
  #           by = c("cruise_id",
  #                  "site" = "locationID",
  #                  "mesh" #,
  #                  # "recordedBy",
  #                  #"recordedByID", "basisOfRecord"
  #                  )
  #           ) %>%
  # dplyr::transmute(
  # this method is prefered to maintainers instead of transmute
  dplyr::mutate(
    .keep = "none",
    # ---

    decimalLatitude,
    decimalLongitude,
    # eventID,
    # eventDate                = glue("{date}T{local_time_est}ET"),
    # make datetime and convert to utc
    eventDate = lubridate::ymd_hms(
      glue::glue("{date} {local_time_est}",
        .na = ""
      ),
      tz = "EST",
      truncated = 3
    ),
    eventDate = lubridate::format_ISO8601(eventDate, usetz = "Z"),
    # eventDate = if_else(is.na(eventDate), as.character(date), eventDate),

    # occurrenceID             = glue("{eventID}:{aphiaID}:{lifeStage}"),
    # changed to taxa_orig to make sure occur ID is unique
    # occurrenceID             = glue("{eventDate}:{eventID}:{taxa_orig}:{lifeStage}"),
    occurrenceID = glue("{datasetID}:{sheet_nm}:{clasification}"),
    occurrenceID = str_replace_all(occurrenceID, " ", "_"),
    # taxa names
    scientificName,
    # across(kingdom:genus),
    # taxonRank = rank,
    # taxonID = taxon_rank_id,

    # recorded by
    # recordedBy,
    # recordedByID,
    # dateIdentified           = date_analyzed,

    # counts of organisms?
    # IDk which of these is accurate
    # individualCount          = number_ind_sample,
    # add each aliquot as a total count per sample event
    individualCount,
    # organismQuantity         = ind_m3,
    # organismQuantityType     = "Individuals per cubic metre",
    # or
    # measurementType          = "Number per cubic metre",
    # MOF? should also include ind/sample?
    # measurementUnitID        =
    # "http://vocab.nerc.ac.uk/collection/P06/current/UPMM/",
    # MOF?

    # TODO: check this information ----
    lifeStage,
    # establishmentMeans       = "native | uncertain",
    occurrenceStatus = "present",
    preparations = "formalin before analysis | ethanol after analysis",
    scientificNameID,
    # like urn:lsid:ipni.org:names:37829-1:1.3
    basisOfRecord = "PreservedSpecimen", # "humanObservation",
    datasetID = {
      datasetID
    },
    identificationReferences = "WoRMS",
    verbatimIdentification = clasification,
    # georeferenceVerificationStatusProperty = "verified by contributor",
    georeferenceVerificationStatus = "verified by contributor",
    dispostion = "in collection",
    coordinateUncertaintyInMeters = 500,
    minimumDepthInMeters = 0,
    maximumDepthInMeters,
    # identifiedBy = "Jamie"
  )


filter(occur, is.na(scientificName)) %>%
 count(verbatimIdentification) 

occur <- filter(occur, !is.na(scientificName))
naniar::vis_miss(occur)
# naniar::vis_miss(dat_merg_tx)
# naniar::vis_miss(dat_merg)
Hmisc::describe(occur)
```


```{r check-dupes}
janitor::get_dupes(occur, occurrenceID)
```

```{r construct-occurrence-table}
# save locally
if (FALSE) {
  readr::write_excel_csv(
    occur,
    here::here(
      base_dir, "processed",
      glue(
        "occurrences",
        "{format(Sys.time(), '_%Y%m%d_%H%M%S')}.", "csv"
      )
    )
  )
}


# save to cloud if connection exists
if (FALSE && exists("cloud_dir") && !is.na(cloud_dir)) {
  master_taxa_list(
    taxa_list  = occur,
    .cloud_dir = here(cloud_dir, "old_jaime_data"),
    where_to   = "cloud",
    save       = TRUE,
    .file_expr = "occurrences",
    sheet_name = "Simplified Occurence Core"
  )
}
```

# 5.0 ---- Add More Cores ----
Quick first test.
TODO: update as needed

## 5.1 Record Level
```{r rcd-lvl}
taxa_matched_merg <-
  dat_merg_tx %>%
  mutate(
    type = "Event",
    modified = Sys.Date(),
    language = "en",
    license = "http://creativecommons.org/publicdomain/zero/1.0/legalcode",
    # institutionCode     = "USF_IMaRS",
    institutionCode = "USF",
    # parentEventID       = "IMaRS_MBON_zooplankton",
    datasetID = "USF_IMaRS_MBON_compiled_zoo_taxonomy_jaimie_2018",
    datasetName = paste(
      "MBON/USF-IMaRS Florida Keys National Marine Sanctuary",
      "Zooplankton Net Tows (2015 - 2017)"
    ),
    basisOfRecord = "PreservedSpecimen", # "HumanObservation",
    # informationWithheld = "collector identities withheld because changed frequently",
    # recordedBy          = nlf,
    recordedBy = "NOAA AOML",
    # recordedByID        = orcid_nat,
    # Abbey B. recommends this change, updated March 21, 2023
    identifiedBy = "Jamie <TODO>",
    tow_time = case_when(
      str_detect(tow_time_min, ":") ~ str_extract(tow_time_min, "(\\d+):", 1),
      .default = tow_time_min
    ),
    tow_time = parse_number(tow_time),
    tow_time = round(tow_time, 2)
  )

rcd_lvl <- taxa_matched_merg %>%
    distinct()
```

## 5.2 Event Level Info
```{r event}
event <-
    taxa_matched_merg  %>%
    mutate(
        .keep = "none",
        # cruise_id, 
        sheet_nm,
        mesh, datasetID, type, modified, license, 
        institutionCode, datasetName, 
        language,
        locationID         = station,
        eventDate          = lubridate::ymd_hms(
            stringr::str_c(date, local_time_est),
            tz = "EST"),
        eventDate          = lubridate::format_ISO8601(eventDate, usetz = "Z"),
        eventDate          = if_else(is.na(eventDate), as.character(date), eventDate),
        eventID            = glue("{datasetID}:{sheet_nm}"),
        year               = year(date),
        month              = month(date),
        day                = day(date),
        samplingProtocol   = glue(
            "{mesh} mesh size (um) - ",
            "bongo nets | folson splitter | http://drs.nio.org/drs/handle/2264/95"
        ),
        
        # TODO: habitat NERC vocab
        habitat            = "near reef",
        sampleSizeValue    = vol_filtered_m_3,
        sampleSizeUnit     = "Volume per cubic metre of filtered seawater",
        samplingEffort     = str_c(tow_time, "minutes", sep = " ")
    ) %>%
    distinct() %>%
    mutate(
        # TODO: save for the end when combining all same sizes together
        catalogNumber   = row_number(),
        .before = everything()
     ) 
```

## 5.3 Location Level Info:
```{r location}
location <-
    taxa_matched_merg %>%
    left_join(event, 
              by = join_by("sheet_nm", "station" == "locationID", "mesh")) %>%
    dplyr::mutate(
        .keep = "none",
        eventID,
        decimalLatitude, 
        decimalLongitude,
        higherGeographyID = case_when(
            str_detect(stn, "MR|LK|WS|9B") ~ "http://vocab.getty.edu/tgn/7030258",
            str_detect(stn, "5") ~ "http://vocab.getty.edu/tgn/1101513",
            TRUE ~ "Not Found"
        ),
        higherGeography   = "North America | United States | Florida",
        continent         = "North America",
        country           = "United States",
        countryCode       = "US",
        stateProvince     = "Florida",
        geodeticDatum     = "EPSG:4326",
        georeferencedBy   = "NOAA AOML | USF IMaRS | RSMAS R/V Walton Smith",
        coordinateUncertaintyInMeters = 500,
        minimumDepthInMeters = 0,
        maximumDepthInMeters
    ) %>%
    distinct()

left_join(
    event, location
) 
```

## 5.4 Ocurrence Level Info:
```{r occurence}
occur <-
  taxa_matched_merg %>%
  left_join(event,
    by = join_by(
      "sheet_nm",
      "station" == "locationID",
      "mesh"
    )
  ) %>%
    dplyr::mutate(
        .keep = "none",
        eventID, 
        occurrenceID             = glue("{eventID}:{clasification}"),
        occurrenceID             = str_replace_all(occurrenceID, " ", "_"),
        
        # taxa names
        scientificName,
        across(kingdom:genus), # TODO: check if species column exists
        taxonRank = rank,
        organismQuantity         = individualCount,
        organismQuantityType     = "Summation of 1 mL Aliquots",
        

        lifeStage,
        occurrenceStatus         = "present",
        preparations             = "formalin before analysis | ethanol after analysis",
        scientificNameID,
        basisOfRecord,
        identificationReferences = "WoRMS",
        verbatimIdentification   = clasification,
        georeferenceVerificationStatus = "verified by contributor",
        dispostion = "in collection"
    ) %>%
      relocate(eventID, occurrenceID, 
               .before = 1) %>%
    
    # removes row where couldn't get accurate aphia ID
    filter(!is.na(scientificNameID))


```

# 5.5 Examples to save
```{r final-to-core}
event_final <- 
    left_join(event, location) %>%
    select(-sheet_nm, -datasetID, -catalogNumber, -mesh) 

# visualize missing    
naniar::vis_miss(event_final)
naniar::vis_miss(occur)

# how to merge event and occur
full_join(event_final, occur, by = "eventID") %>%
    naniar::vis_miss()

# check duplicates
stopifnot(nrow(janitor::get_dupes(event_final, eventID)) == 0)
stopifnot(nrow(janitor::get_dupes(occur, occurrenceID)) == 0) 
```


```{r save}
# save obis cores locally
if (FALSE) {
  dir_sv <- here(base_dir, "obis")
  path <-
    here(
      dir_sv,
      glue(
        "{c('event', 'occur', 'mof')}",
        "_example_update_{Sys.Date()}.csv"
      )
    )
  path

  dir_create(dir_sv)

  event_final %>%
    write_excel_csv(file = path[1], na = "")

  occur %>%
    write_excel_csv(file = path[2], na = "")
}

# save obis cores to cloud if connection exists
if (FALSE && exists("cloud_dir") && !is.na(cloud_dir)) {

    master_taxa_list(
    taxa_list  = event_final,
    .cloud_dir = here(cloud_dir, "old_jaime_data", "obis_cores"),
    where_to   = "cloud",
    save       = TRUE,
    .file_expr = "event_example",
    sheet_name = "Event Core"
  )   
    master_taxa_list(
    taxa_list  = occur,
    .cloud_dir = here(cloud_dir, "old_jaime_data", "obis_cores"),
    where_to   = "cloud",
    save       = TRUE,
    .file_expr = "occur_example",
    sheet_name = "Occurence Core"
  )   
}
```