---
title: "Fix and Convert Previous Zooplankton Data"
author: "Sebastian Di Geronimo"
date: "2023-04-09"
format: html
---

# desc
Inputs the raw older (Jaime's) files, outputs a `.csv` file with all data & metadata in one table.

## old file notes
The methodology allows you to ignore common species and focus on the uncommmon species.
The protocol is to count the common species `N_(A=1)` and assume that the same number are in later counts.
Rarer species are counted in each aliquot.

* `Total Ind. Sample` is the calculated amount of individuals in the total sample from the net.
* actual counted individuals under the microsocope is the number of individuals in the 1mL taken and diluted 
* column B is the manual counted number of individuals
* column C is column B * `Counted aliquot` (`F7`)
    * WS03165002 sheet does not follow this rule
    * column C is meant to total be the total number of individuals 
        * sometimes it is fractional; that makes no sense.
            * eg `LK111564`
        
## new file notes

* `total_ind_sample` *should* be the number of individuals of that species which were identified. This is *actually* the density of individuals per milliliter.
* `n_ind_m3` 
* `ind_count` is the number of individuals identified under the microscope. This is what is used as the count for OBIS.

The output file should have everything we need but some cruises are missing sample details information such as:
* vol_filtered_m3

Additional columns are there too.

* no_ind_aliquot

Jaime counted a certain number of a species, if it reached a certain amount:
* counted_aliquot : this is multipled by number of aliquots sometimes to get the total individual count
* 


# Load Libraries
```{r setup}
librarian::shelf(
    librarian, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    # broom # optional
    
    # additional
    readxl, janitor, hablar
)

library("conflicted")

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```
```{r create-folder}
base_dir <- here("data", "zoo_pre_2018")

dir_create(here(base_dir, "processed"))
dir_create(here(base_dir, "orig"))
```

# Search for Files
two files needed:
compiled_zoo_taxonomy_Jaimie_31JAN2018 (1).xlsx
sample-details.csv

Download these and put in the origins folder
https://docs.google.com/spreadsheets/d/1GmebUJ35rlVF07b2GNWZXLj91RlDc5lW/edit#gid=1910593077
https://drive.google.com/drive/folders/17f5gOJuU6wBuoNvkpoc7fkW4-8Ql6B3e

```{r get-sheet-info}
file_location <-
    base_dir %>%
    dir_ls(regexp = "[^~].compiled_zoo",
           recurse = TRUE)

meta_samples_file <-
    base_dir %>%
    dir_ls(regexp = "[^~].sample-details",
           recurse = TRUE)

file_location
meta_samples_file

# read all sheet names and filter for ones with WS and LK
# WS = Western Sambo
# LK = Looe Key
sheets <-
    excel_sheets(file_location) %>%
    
    # turn into tibbles (DON'T worry about what it is)
    tibble(sheet_nm = .) %>%
    
    # filter(str_detect(sheets, "^(WS|LK)")) %>%
    filter(!str_detect(sheet_nm, "Samples|PL")) %>%
    mutate(
        sheet = if_else(str_detect(sheet_nm, "MR116500"), "MR0116500", sheet_nm),
        sheet = if_else(str_detect(sheet_nm, "WS0117640"), "WS011764", sheet)) %>%
    separate_wider_regex(
        sheet,
        cols_remove = TRUE,
        patterns = c(stn   = "\\w{2}",      # extract station
                     month = "\\d{2}",      # extract month
                     year  = "\\d{2}",      # extract year
                     mesh  = "200|500|64",  # extract mesh
                     dup   = ".*|2$"))  %>% # extract duplicates
    arrange(year, month, stn, as.numeric(mesh)) %>%
    mutate(dup   = if_else(str_detect(dup, "2"), 
                           "duplicate", 
                           NA_character_),
           year = str_c("20", year)) %>%
    relocate(stn, .after = year) %>%
    relocate(year, .after = 1)
```

# Optional: Reorder Sheets
When opening in excel, it is hard to search through each sheet because they are
not in a year, month, station, mesh order. This chunk will fix this issue, but 
the created file will need `repairs` before using. This is only used when 
looking for issues in the raw data and is not used elsewhere. 
```{r reorder-sheets}
if (!file_exists(here(base_dir, "processed", "zoo_reorder_sht.xlsx"))
    & FALSE) {
    shelf(openxlsx)

    wb <- loadWorkbook(file_location)

    reorder <-
        c(
          # first two sheets
          (which(str_detect(names(wb), "Samples for see|PLANTILLA"))),
          
          # reordered sheets
          (names(wb) %>%
          tibble(file = .) %>%
          mutate(rows = row_number()) %>%
          left_join(sheets, 
                    ., 
                    by = join_by("sheet_nm" == "file")) %>%
          pull(rows))
          )
    
    # reordering sheets
    openxlsx::worksheetOrder(wb) <- reorder
    
    # saving sheets
    openxlsx::saveWorkbook(
        wb, 
        here(base_dir, "zoo_reorder_sht.xlsx"),
        overwrite = TRUE)
    
    unshelf(openxlsx)
    rm(wb, reorder)
} else {
    message("Not creating reordered xlxs sheet file.")
}
```

# Read Zooplankton Data
```{r read-data}
# read each sheet for data and metadata
species <- 
    sheets %>%
    
    # add 2 columns and loop through each file to load
    # 1. meta data (includes date and location)
    # 2. data (includes species info)
    mutate(
        # metadata
        meta = map(sheet_nm, 
                   ~read_xlsx(file_location, 
                              sheet = .,
                              n_max = 7, 
                              col_names    = FALSE,
                              .name_repair = make_clean_names) 
                   ),
        # extract important metadata
        meta_top = map(meta,
                       function(x) {
                           r <- which(str_detect(x$x, "Vol|^\\.$"))
                           
                           if (any(str_detect(x$x, "^\\.$"), 
                                   na.rm = TRUE)) 
                               x$x[r] <- "Vol filtered  (m³ )"
                           
                           x_name <-
                               slice(x, r) %>%
                               unlist(use.names = FALSE) %>%
                               make_clean_names()
                           x <- slice(x, r + 1)
                           names(x) <- x_name
                           x <- remove_empty(x, which = "cols")
                           }
                       ),
        # species data
        data = map2(sheet_nm, 
                    meta_top,
                   ~read_xlsx(file_location, 
                              sheet = .x, 
                              skip  = 7,
                              na    = c("", "#VALUE!"),
                              .name_repair = make_clean_names
                              )  %>%
                     filter(!str_detect(clasification, "Densidad total")) %>%
                     rename("ind_count" = x) %>%
                     bind_cols(.y) %>%
                     mutate(
                        vol_filtered_m3 = if_else(!str_detect(vol_filtered_m3, 
                                                       "no hay volumen|XXX"),
                                           vol_filtered_m3, 
                                           NA_character_) %>%
                                           as.numeric()
                     ) %>%
                     hablar::retype()
                   )
    ) %>%
    select(-meta_top)

species_unnest <- 
    species %>%
    
    # extract the data within the list
    unnest(data)
```


```{r testing-data}
species_unnest 
species_unnest  %$% unique(vol_filtered_m3)
# species_unnest  %>%
#     filter(if_any(contains("vol_filt"):counted_aliquot, 
#                   (\(x) is.na(x) | str_detect(x, "no hay volumen|XXX"))
#                   )) 
# 
# shell.exec(file_location)
```

# Report No Volume Recorded
```{r}
# find NAs in n_ind_m3 and count sheet
vol_miss <- 
    species_unnest %>%
    filter(is.na(n_ind_m3)) %>%
    count(sheet_nm) %$% 
    unique(sheet_nm)

# determine if na for individuals per cubic meter are because:
# 1. there was no volume recorded
# 2. the species didn't exist
species1 <-
  species %>%
  mutate(
    data = map2(
      .x   = sheet_nm,
      .y   = data,
      vols = vol_miss,
      function(x, y, vols) {
        mutate(
          y,
          no_vol = if_else(
            x %in% vols,
            "no volume recorded", NA_character_
          )
        )
      }
    )
  )

rm(vol_miss)
```
# Merge Meta Data with Zooplankton Data
```{r}
# load metadata and fix
meta_samples <-
  meta_samples_file %>%
  read_csv(
    show_col_types = FALSE,
    name_repair = make_clean_names
  ) %>%
  select(-x) %>%
  filter(!if_all(2:last_col(), \(x) is.na(x))) %>%
  mutate(
    month   = format(date, "%m"),
    year    = format(date, "%Y"),
    day     = format(date, "%d"),
    .before = date
  ) %>%
  mutate(
    stn = case_when(
      str_detect(station, "Mol")  ~ "MR",
      str_detect(station, "West") ~ "WS",
      str_detect(station, "Loo")  ~ "LK",
      .default = station
    ),
    .before = mesh_size_um
  ) %>%
  separate_longer_delim(mesh_size_um, delim = " / ") %>%
  mutate(
    lat_in = suppressWarnings(parzer::parse_lat(lat_in)), # warns if NA
    lon_in = suppressWarnings(parzer::parse_lon(lon_in)), # warns if NA
    lat_in = if_else(is.numeric(lat_in) & lat_in < 0, -lat_in, lat_in),
    lon_in = if_else(is.numeric(lon_in) & lon_in > 0, -lon_in, lon_in)
  ) %>% 
  nest(
      .by = c(year,
            month,
            stn,
            mesh_size_um)
  )
    # get_dupes(stn, date, mesh_size_um)
    # distinct(stn, Date, `mesh size (um)`, .keep_all = TRUE)



# merge metadata with zoo data
# this combines month, year, station and mesh 
species1 <-
    species %>%
    # left_join(
    full_join(
        meta_samples,
        by = join_by(
            year,
            month,
            stn,
            "mesh"  == "mesh_size_um")
    ) %>%
      clean_names()

species3 <-
    species1 %>%
    unnest(data_x, names_repair = make_clean_names) %>% 
    unnest(data_y, names_repair = make_clean_names) %>%
    remove_empty("cols") %>%
  mutate(
    split_size_2 = map_dbl(
      split_size_2,
      function(x) {
        out <- tryCatch(
          {
            eval(parse(text = x))
          },
          error = function(e) {
            NA_integer_
          }
        )
      }
    )
  ) %>%
    relocate(contains("clasific"):n_ind_m3, 
             .after = last_col()) %>%
 relocate(day, date, local_time_est, station, 
          .after = month)              
```

# Save Files
```{r save-files}
if (FALSE) {
    write_csv(
        species,
        file = here(base_dir, "processed",
                    glue("zoo_compiled",
                         format(Sys.Date(), "_%Y%m%d"),
                         ".csv")),
        na = "")
    }
if (FALSE) {
    species3 %>%
    select(-meta) %>%
    write_csv(,
              file = here(base_dir, "processed",
                          glue("zoo_compiled_with_meta",
                               format(Sys.Date(), "_%Y%m%d"),
                               ".csv")),
              na   = "")
}
```

# Read 2015 - 2017 OBIS Files (needs fixing)
```{r}

"https://ipt-obis.gbif.us/archive.do?r=sfmbon_zooplankton&v=1.5"
sf_mbon <- 
    here("data", "example", 
         "claudia_zoo_biometry",
         "dwca-sfmbon_zooplankton-v1.5", "occurrence.txt") %>%
    read_delim(
        delim = "\t", 
        escape_double = FALSE, 
        trim_ws = TRUE,
        show_col_types = FALSE)

sf_mbon %>%
    filter(
        !eventDate > as.Date("2017-01-01")
    ) %>%
      View()

sf_mbon %$% unique(eventDate) %>%
    as.Date() %>%
    sort()

here("data", "example", "claudia_zoo_biometry", 
     "0122465-220831081235567", "verbatim.txt") %>%
    read_delim(.,
               delim = "\t", 
               escape_double = FALSE,
               trim_ws = TRUE) %>%
    remove_empty(which = "cols") %>%
    View()

here("data", "example", "claudia_zoo_biometry", 
     "0122465-220831081235567", "occurrence.txt") %>%
    read_delim(.,
               delim = "\t", 
               escape_double = FALSE,
               trim_ws = TRUE,
               show_col_types = FALSE) %>%
    remove_empty(which = "cols") %>%
    select(1:day) %>%
    View()



```

