---
title: "Fix and Convert Previous Zooplankton Data"
author: "Sebastian Di Geronimo"
date: "2023-04-09"
format: html
---
# Load Libraries
```{r setup}
librarian::shelf(
    librarian, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here,
    # broom # optional
    
    # additional
    readxl, janitor, hablar
)

library("conflicted")

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```
# Search for Files
two files needed:
compiled_zoo_taxonomy_Jaimie_31JAN2018 (1).xlsx
sample-details.csv

```{r get-sheet-info}
base_dir <- here("data", "example", "claudia_zoo_biometry")

file_location <-
    base_dir %>%
    dir_ls(regexp = "[^~].compiled_zoo")

meta_samples_file <-
    base_dir %>%
    dir_ls(regexp = "[^~].sample-details")

file_location
meta_samples_file

# read all sheet names and filter for ones with WS and LK
# WS = Western Sambo
# LK = Looe Key
sheets <-
    excel_sheets(file_location) %>%
    
    # turn into tibbles (DON'T worry about what it is)
    tibble(sheet_nm = .) %>%
    
    # filter(str_detect(sheets, "^(WS|LK)")) %>%
    filter(!str_detect(sheet_nm, "Samples|PL")) %>%
    mutate(
        sheet = if_else(str_detect(sheet_nm, "MR116500"), "MR0116500", sheet_nm),
        sheet = if_else(str_detect(sheet_nm, "WS0117640"), "WS011764", sheet)) %>%
    separate_wider_regex(
        sheet,
        cols_remove = TRUE,
        patterns = c(stn   = "\\w{2}",      # extract station
                     month = "\\d{2}",      # extract month
                     year  = "\\d{2}",      # extract year
                     mesh  = "200|500|64",  # extract mesh
                     dup   = ".*|2$"))  %>% # extract duplicates
    arrange(year, month, stn, as.numeric(mesh)) %>%
    mutate(dup   = if_else(str_detect(dup, "2"), 
                           "duplicate", 
                           NA_character_),
           year = str_c("20", year)) %>%
    relocate(stn, .after = year) %>%
    relocate(year, .after = 1)
```

# Optional: Reorder Sheets
When opening in excel, it is hard to search through each sheet because they are
not in a year, month, station, mesh order. This chunk will fix this issue, but 
the created file will need `repairs` before using. This is only used when 
looking for issues in the raw data and is not used elsewhere. 
```{r reorder-sheets}
if (FALSE) {
    shelf(openxlsx)

    wb <- loadWorkbook(file_location)

    reorder <-
        c(
          # first two sheets
          (which(str_detect(names(wb), "Samples for see|PLANTILLA"))),
          
          # reordered sheets
          (names(wb) %>%
          tibble(file = .) %>%
          mutate(rows = row_number()) %>%
          left_join(sheets, 
                    ., 
                    by = join_by("sheet_nm" == "file")) %>%
          pull(rows))
          )
    
    # reordering sheets
    openxlsx::worksheetOrder(wb) <- reorder
    
    # saving sheets
    openxlsx::saveWorkbook(
        wb, 
        here(base_dir, "zoo_reorder_sht.xlsx"),
        overwrite = TRUE)
    
    unshelf(openxlsx)
    rm(wb, reorder)
}
```

# Read Zooplankton Data
```{r read-data}
# read each sheet for data and metadata
species <- 
    sheets %>%
    
    # add 2 columns and loop through each file to load
    # 1. meta data (includes date and location)
    # 2. data (includes species info)
    mutate(
        # metadata
        meta = map(sheet_nm, 
                   ~read_xlsx(file_location, 
                              sheet = .,
                              n_max = 7, 
                              col_names    = FALSE,
                              .name_repair = make_clean_names) 
                   ),
        # extract important metadata
        meta_top = map(meta,
                       function(x) {
                           r <- which(str_detect(x$x, "Vol|^\\.$"))
                           
                           if (any(str_detect(x$x, "^\\.$"), 
                                   na.rm = TRUE)) 
                               x$x[r] <- "Vol filtered  (m³ )"
                           
                           x_name <-
                               slice(x, r) %>%
                               unlist(use.names = FALSE) %>%
                               make_clean_names()
                           x <- slice(x, r + 1)
                           names(x) <- x_name
                           x <- remove_empty(x, which = "cols")
                           }
                       ),
        # species data
        data = map2(sheet_nm, 
                    meta_top,
                   ~read_xlsx(file_location, 
                              sheet = .x, 
                              skip  = 7,
                              na    = c("", "#VALUE!"),
                              .name_repair = make_clean_names
                              )  %>%
                     filter(!str_detect(clasification, "Densidad total")) %>%
                     rename("ind_count" = x) %>%
                     bind_cols(.y) %>%
                     mutate(
                        vol_filtered_m3 = if_else(!str_detect(vol_filtered_m3, 
                                                       "no hay volumen|XXX"),
                                           vol_filtered_m3, 
                                           NA_character_) %>%
                                           as.numeric()
                     ) %>%
                     hablar::retype()
                   )
    ) %>%
    select(-meta_top)

species_unnest <- 
    species %>%
    
    # extract the data within the list
    unnest(data) 

species_unnest 
species_unnest  %$% unique(vol_filtered_m3)
# species_unnest  %>%
#     filter(if_any(contains("vol_filt"):counted_aliquot, 
#                   (\(x) is.na(x) | str_detect(x, "no hay volumen|XXX"))
#                   )) 
# 
# shell.exec(file_location)
```

# Report No Volume Recorded
```{r}
# find NAs in n_ind_m3 and count sheet
vol_miss <- 
    species_unnest %>%
    filter(is.na(n_ind_m3)) %>%
    count(sheet_nm) %$% 
    unique(sheet_nm)

# determine if na for individuals per cubic meter are because:
# 1. there was no volume recorded
# 2. the species didn't exist
species1 <-
  species %>%
  mutate(
    data = map2(
      .x   = sheet_nm,
      .y   = data,
      vols = vol_miss,
      function(x, y, vols) {
        mutate(
          y,
          no_vol = if_else(
            x %in% vols,
            "no volume recorded", NA_character_
          )
        )
      }
    )
  )

rm(vol_miss)
```
# Merge Meta Data with Zooplankton Data
```{r}
# load metadata and fix
meta_samples <-
  meta_samples_file %>%
  read_csv(
    show_col_types = FALSE,
    name_repair = make_clean_names
  ) %>%
  filter(!if_all(2:last_col(), \(x) is.na(x))) %>%
  mutate(
    month   = format(date, "%m"),
    year    = format(date, "%Y"),
    day     = format(date, "%d"),
    .before = date
  ) %>%
  mutate(
    stn = case_when(
      str_detect(station, "Mol")  ~ "MR",
      str_detect(station, "West") ~ "WS",
      str_detect(station, "Loo")  ~ "LK",
      .default = station
    ),
    .before = mesh_size_um
  ) %>%
  separate_longer_delim(mesh_size_um, delim = " / ") %>%
  mutate(
    lat_in = suppressWarnings(parzer::parse_lat(lat_in)), # warns if NA
    lon_in = suppressWarnings(parzer::parse_lon(lon_in)), # warns if NA
    lat_in = if_else(is.numeric(lat_in) & lat_in < 0, -lat_in, lat_in),
    lon_in = if_else(is.numeric(lon_in) & lon_in > 0, -lon_in, lon_in)
  ) %>% 
  nest(
      .by = c(year,
            month,
            stn,
            mesh_size_um)
  )
    # get_dupes(stn, date, mesh_size_um)
    # distinct(stn, Date, `mesh size (um)`, .keep_all = TRUE)



# merge metadata with zoo data
# this combines month, year, station and mesh 
species1 <-
    species %>%
    # left_join(
    full_join(
        meta_samples,
        by = join_by(
            year,
            month,
            stn,
            "mesh"  == "mesh_size_um")
    ) 

species3 <- 
    species1 %>%
    unnest(data.x, names_repair = "unique") %>%
    unnest(data.y, names_repair = "unique") %>%
    remove_empty("cols") %>%
  mutate(
    split_size...36 = map_dbl(
      split_size...36,
      function(x) {
        out <- tryCatch(
          {
            eval(parse(text = x))
          },
          error = function(e) {
            NA_integer_
          }
        )
      }
    )
  )
```

# Save Files
```{r save-files}
if (FALSE) {
    write_csv(
        species,
        file = here(base_dir, 
                    glue("zoo_compiled",
                         format(Sys.Date(), "_%Y%m%d"),
                         ".csv")),
        na = "")
    }
if (FALSE) {
    species3 %>%
    select(-meta)
    write_csv(,
              file = here(base_dir, 
                          glue("zoo_compiled_with_meta",
                               format(Sys.Date(), "_%Y%m%d"),
                               ".csv")),
              na   = "")
}
```

# Read 2015 - 2017 OBIS Files (needs fixing)
```{r}

"https://ipt-obis.gbif.us/archive.do?r=sfmbon_zooplankton&v=1.5"
sf_mbon <- 
    here("data", "example", 
         "claudia_zoo_biometry",
         "dwca-sfmbon_zooplankton-v1.5", "occurrence.txt") %>%
    read_delim(
        delim = "\t", 
        escape_double = FALSE, 
        trim_ws = TRUE,
        show_col_types = FALSE)

sf_mbon %>%
    filter(
        !eventDate > as.Date("2017-01-01")
    ) %>%
      View()

sf_mbon %$% unique(eventDate) %>%
    as.Date() %>%
    sort()

here("data", "example", "claudia_zoo_biometry", 
     "0122465-220831081235567", "verbatim.txt") %>%
    read_delim(.,
               delim = "\t", 
               escape_double = FALSE,
               trim_ws = TRUE) %>%
    remove_empty(which = "cols") %>%
    View()

here("data", "example", "claudia_zoo_biometry", 
     "0122465-220831081235567", "occurrence.txt") %>%
    read_delim(.,
               delim = "\t", 
               escape_double = FALSE,
               trim_ws = TRUE,
               show_col_types = FALSE) %>%
    remove_empty(which = "cols") %>%
    select(1:day) %>%
    View()



```

