---
title: "Fix and Convert Previous Zooplankton Data"
author: "Sebastian Di Geronimo"
date: "2023-04-09"
format: html
---

# desc
Inputs the raw older (Jaime's) files, outputs a `.csv` file with all data & metadata in one table.

## old file notes
The methodology allows you to ignore common species and focus on the uncommmon species.
The protocol is to count the common species `N_(A=1)` and assume that the same number are in later counts.
Rarer species are counted in each aliquot.

* `Total Ind. Sample` is the calculated amount of individuals in the total sample from the net.
* actual counted individuals under the microsocope is the number of individuals in the 1mL taken and diluted 
* column B is the manual counted number of individuals
* column C is column B * `Counted aliquot` (`F7`)
    * WS03165002 sheet does not follow this rule
    * column C is meant to total be the total number of individuals 
        * sometimes it is fractional; that makes no sense.
            * eg `LK111564`
        
## new file notes

* `total_ind_sample` *should* be the number of individuals of that species which were identified. This is *actually* the density of individuals per milliliter.
* `n_ind_m3` 
* `ind_count` is the number of individuals identified under the microscope. This is what is used as the count for OBIS.

The output file should have everything we need but some cruises are missing sample details information such as:
* vol_filtered_m3

Additional columns are there too.

* no_ind_aliquot

Jaime counted a certain number of a species, if it reached a certain amount:
* counted_aliquot : this is multipled by number of aliquots sometimes to get the total individual count
* 


# Load Libraries
```{r setup}
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
  quiet = TRUE,
  librarian, conflicted, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
  forcats, lubridate, glue, fs, magrittr, here,
    
  # additional
  readxl,
  # xlsx,
  janitor, hablar,
    
)

conflicts_prefer(
  dplyr::filter(),
  dplyr::select()
)
```

```{r create-folder}
base_dir <- here("data", "zoo_pre_2018")

dir_create(here(base_dir, "processed"))
dir_create(here(base_dir, "orig"))
```

# Search for Files
two files needed:
compiled_zoo_taxonomy_Jaimie_31JAN2018 (1).xlsx
sample-details.csv

Download these and put in the orig folder
https://docs.google.com/spreadsheets/d/1GmebUJ35rlVF07b2GNWZXLj91RlDc5lW/edit#gid=1910593077
https://drive.google.com/drive/folders/17f5gOJuU6wBuoNvkpoc7fkW4-8Ql6B3e

```{r file-search}
file_location <-
  base_dir %>%
  dir_ls(
    regexp = "[^~].compiled_zoo|zoo_reorder_sht",
    recurse = TRUE)  %>%
  str_subset("~\\$", negate = TRUE) %T>%
  print() 

# select reorder file if exists, will be created lower
file_location <-
  if_else(
    !is_empty(str_subset(file_location, "reorder")),
    str_subset(file_location, "reorder"),
    str_subset(file_location, "compiled")
    ) %T>%
  print() 

# get metadata from collection
meta_samples_file <-
  base_dir %>%
  dir_ls(
    regexp = "[^~].sample-details",
    recurse = TRUE) %T>%
  print()


```

# Read all sheet names and filter for ones with MR, WS and LK

MR = Molasses Reef
WS = Western Sambo
LK = Looe Key

```{r get-sheet-info}
sheets <-
  readxl::excel_sheets(file_location) %>%
  tibble(sheet_nm = .) %>%
  filter(!str_detect(sheet_nm, "Samples|PL")) %>%
  mutate(
    # may need to change to this?
    sheet = if_else(str_detect(sheet_nm, "MR116500"), "MR1116500", sheet_nm),
    # sheet = if_else(str_detect(sheet_nm, "WS0117640"), "WS011764", sheet)
  ) %>%
  separate_wider_regex(
    sheet,
    cols_remove = TRUE,
    patterns    = c(
      stn = "\\w{2}", # extract station
      month = "\\d{2}", # extract month
      year = "\\d{2}", # extract year
      mesh = "200|500|64", # extract mesh
      dup = ".*|0$|2$"
    )
  ) %>% 
  # extract duplicates
  arrange(year, month, stn, mesh) %>%
  mutate(
    year = str_c("20", year),
    dup = if_else(str_detect(dup, "0|2"),
      "duplicate",
      NA_character_
    )
  ) %>%
  relocate(stn, .after = year) %>%
  relocate(year, .after = 1) %T>%
  print()

arrange(sheets, sheet_nm) %>%
  mutate(
    sheet_num2 =
      case_when(
        !is.na(dup) ~ str_remove(sheet_nm, "(0|2)$"),
        .default = sheet_nm
      )
  ) %>%
  janitor::get_dupes(sheet_num2)
```

# Determine Differences of Individuals per Sample

```{r}
if (FALSE) {
test <-
  sheets %>%
  # slice(c(1, 5, 34)) %>%
  mutate(
    # metadata
    meta = imap(
      sheet_nm,
      \(.x, idx) {
        print(idx)
        
        shelf(xlsx)
        
        dat <- xlsx::read.xlsx(
          file         = file_location,
          sheetIndex   = idx + 2,
          startRow     = 8,
          colIndex     = c(4, 6),
          keepFormulas = TRUE
        ) %>%
       janitor::remove_empty("rows") %>%
       mutate(
            `Total.Ind..Sample` = as.character(`Total.Ind..Sample`)
          )

        unshelf(xlsx)
        
        return(dat)
        }
    ),
  ) 

test2 <- 
  test %>%
  # extract the data within the list
  unnest(meta)  %>%
  rename("D" = Total.Ind..Sample) %T>%
  print()

test2 

test_calc <-
  test2 %>%
  mutate(
    .before = 1,
    .by = sheet_nm,
    row = row_number()
  ) %>%
  mutate(
    calc =
      case_when(
        str_detect(`D`, "\\$C\\$7") ~ "correct",
        str_detect(`D`, "^\\d.*\\d$") ~ "no calc, only number",
        is.na(`D`) ~ "nothing",
        .default = "no split size"
      ),
    .after = `D`
  )
test_calc %T>%
  print() %T>%
  {
    filter(., str_detect(calc, "nothing")) %>%
    print()
  } 

calc_percent <- 
  test_calc %>%
  filter(!str_detect(calc, "nothing")) %>%
  count(calc) %>%
  mutate(
    frac = n / sum(n) * 100,
    frac = round(frac, 2)
  ) %T>% 
  print()


# list(
#   "calculations" = test_calc,
#   "percents" = calc_percent
# ) %>%
#   writexl::write_xlsx(
#     path = here(base_dir, "formula_issues.xlsx")
#   )
}
```



# Optional: Reorder Sheets
When opening in excel, it is hard to search through each sheet because they are
not in a year, month, station, mesh order. This chunk will fix this issue, but 
the created file will need `repairs` before using. This is only used when 
looking for issues in the raw data and is not used elsewhere. 
```{r reorder-sheets}
if (!file_exists(here(base_dir, "processed", "zoo_reorder_sht.xlsx"))
    & FALSE) {
    shelf(openxlsx)

    wb <- loadWorkbook(file_location)

    reorder <-
        c(
          # first two sheets
          (which(str_detect(names(wb), "Samples for see|PLANTILLA"))),
          
          # reordered sheets
          (names(wb) %>%
          tibble(file = .) %>%
          mutate(rows = row_number()) %>%
          left_join(sheets, 
                    ., 
                    by = join_by("sheet_nm" == "file")) %>%
          pull(rows))
          )
    
    # reordering sheets
    openxlsx::worksheetOrder(wb) <- reorder
    
    # saving sheets
    openxlsx::saveWorkbook(
        wb, 
        here(base_dir, "zoo_reorder_sht.xlsx"),
        overwrite = TRUE)
    
    unshelf(openxlsx)
    rm(wb, reorder)
} else {
    message("Not creating reordered xlxs sheet file.")
}
```

# Read Zooplankton Data

read each sheet for metadata and data

add 2 columns and loop through each file to load:
  1. meta data (includes date and location)
  2. data (includes species info)

Notes on calculations:

`no_ind_aliquot`:
  - if value here, multiply by counted aliquot to get `ind_count`
`ind_count`: 
  - total count of individuals across all aliquots

`total_ind_sample`:
  - count per sample taking into account the number of splits and counted 
    aliquots
  - this is the *issue* because uses excel formula and the majority doesnt 
    add `split_size` to calculation
  - calculation:
    (`total_ind_sample` * `vol_sample_water` / `counted_aliquot`) * 
    (`split_size` / `folson`)
  - however, ~79% of the sheets do not include "* (`split_size` / `folson`)",
    and ~1.6% are only numbers, not calculated
  - 18.53% are calculated correctly
	

```{r read-data}
species <-
  sheets %>%
  mutate(
    # metadata
    meta = imap(
      sheet_nm,
      \(.x, idx) {
        print(idx)
        read_xlsx(
          file_location,
          sheet        = .x,
          n_max        = 7,
          col_names    = FALSE,
          .name_repair = make_clean_names
        )
      }
    ),

    # extract important metadata from row 7 and tidy
    meta_top = map(
      meta,
      function(x) {
        r <- which(str_detect(x$x, "Vol|^\\.$"))

        if (any(str_detect(x$x, "^\\.$"),
          na.rm = TRUE
        )) {
          x$x[r] <- "Vol filtered  (m³ )"
        }

        x_name <-
          slice(x, r) %>%
          unlist(use.names = FALSE) %>%
          make_clean_names()
        x <- slice(x, r + 1)
        names(x) <- x_name
        x <- remove_empty(x, which = "cols")
      }
    ),

    # species data
    data = map2(
      sheet_nm,
      meta_top,
      ~ read_xlsx(file_location,
        sheet = .x,
        skip = 7,
        na = c("", "#VALUE!"),
        .name_repair = make_clean_names
      ) %>%
        filter(!str_detect(clasification, "Densidad total")) %>%
        rename("ind_count" = x)
    )
  )

species_unnest <-
  species %>%
  # extract metadata and data within the list
  unnest(meta_top) %>%

  # convert vol_filtered_m3 to numeric
  mutate(
    .after = vol_filtered_m3,
    vol_filtered_m3 =
      if_else(
        !str_detect(
          vol_filtered_m3,
          "no hay volumen|XXX"
        ),
        vol_filtered_m3,
        NA_character_
      ),
    
    # determine if na for individuals per cubic meter are because:
    # 1. there was no volume recorded
    # 2. the species didn't exist
    no_vol = if_else(is.na(vol_filtered_m3), "no volume recorded", NA_character_)
  ) %>%
  
  hablar::retype() %>%
  unnest(data) %>%
  mutate(
    .after = total_ind_sample,
    total_ind_correct = (ind_count * vol_sample_water / counted_aliquot) * (split_size / folson)
  ) %T>%
  print()

```

# Number of times higher using correct calculation

With correct calculation:

equal = 18.5%
2x higher = 63.8%
4x higher = 15.9% 
different amounts = 1.84%

```{r}
species_unnest %>%
  select(1:dup, folson:counted_aliquot, no_ind_aliquot:total_ind_correct, clasification) %>%
  mutate(
    .after           = total_ind_sample,
    ind_diff         = format(total_ind_correct - total_ind_sample, scientific = FALSE),
    ind_div          = round(total_ind_correct / total_ind_sample, digits = 2),
    total_ind_correct    = format(total_ind_correct, scientific = FALSE),
    total_ind_sample = format(total_ind_sample, scientific = FALSE),
    ind_div_off      = case_when(
      ind_div == 1 ~ "same",
      ind_div == 2 ~ "2x",
      ind_div == 4 ~ "4x",
      ind_div > 4 | ind_div < 0 ~ "large",
      ind_div < 1 ~ "new is less",
      .default = as.character(ind_div)
    )
  ) %T>%
  View("diffs") %>%
  count(ind_div) %>%
  arrange(desc(n)) %>%
  mutate(
    frac = n / sum(n) * 100,
    frac = round(frac, 2)
  )

```


# Merge Meta Data with Zooplankton Data

## Load Metadata

```{r load-meta}
# # load metadata and fix
# meta_samples <-
#   meta_samples_file %>%
#   read_csv(
#     show_col_types = FALSE,
#     name_repair = make_clean_names,
#     na = c("", "NA", "na", "did not work"),
#   ) %>%
#   select(-x) %>%
#   remove_empty("rows") %>%
#   mutate(
#     month   = format(date, "%m"),
#     year    = format(date, "%Y"),
#     day     = format(date, "%d"),
#     .before = date
#   ) %>%
#   fill(month, year, day, date) %>%
#   mutate(
#     stn = case_when(
#       str_detect(station, "Mol") ~ "MR",
#       str_detect(station, "West") ~ "WS",
#       str_detect(station, "Loo") ~ "LK",
#       .default = station
#     ),
#     .before = mesh_size_um
#   ) %>%
#   mutate(
#     lat_in = suppressWarnings(parzer::parse_lat(lat_in)), # warns if NA
#     lon_in = suppressWarnings(parzer::parse_lon(lon_in)), # warns if NA
#     lat_in = if_else( lat_in < 0, -lat_in, lat_in),
#     lon_in = if_else( lon_in > 0, -lon_in, lon_in)
#   ) %>%
#   separate_longer_delim(mesh_size_um, delim = " / ") %>%
#   hablar::retype() %T>%
#   print() %>%
#   nest(
#     .by = c(
#       year,
#       month,
#       stn,
#       mesh_size_um
#     )
#   ) %>%
#   print()


# get_dupes(stn, date, mesh_size_um)
# distinct(stn, Date, `mesh size (um)`, .keep_all = TRUE)
```

```{r load-combine-metadata}
# sheet from IMaRS
meta_df_imars <-
  here(cloud_dir, "cruise_logsheets") %>%
  dir_ls(regexp = "^[^~]*(meta_)+.*\\.csv$") %>%
  last_mod(.) %>%
  read_csv(show_col_types = FALSE) %>%
  filter(date < as_date("2017-04-01")) %T>% 
  print()

# sheet from NOAA
meta_noaa_zoo <- 
  # here(cloud_dir, "cruise_logsheets") %>%
  here("data", "metadata", "cruise_logsheets") %>%
  dir_ls(regexp = "zoo_noaa_meta") %>%
  str_subset("~\\$", negate = TRUE) %>%
  last_mod(.) %>%
  read_csv(show_col_types = FALSE) %>%
  filter(
    year(date_time) == 2015 
    & between(month(date_time), 4, 6)
    & str_detect(locationID, "WS|MR|LK")
  ) %T>% 
  print()

# merged
meta_df <- 
  bind_rows(
    meta_noaa_zoo,
    meta_df_imars
  ) %>%
  select(names(meta_df_imars), everything()) %>%
  mutate(
    month   = format(date, "%m"),
    year    = format(date, "%Y"),
    day     = format(date, "%d"),
    .before = date_time
  )  %>%
  hablar::retype() %>%
  distinct() %T>% 
  print() 

```

## Merge Metadata with Meta

flow out = 112991 - seems a bit high, but that what NOAAs says making vol = 7257.432 m^3

for some reason this was changed in `sample-detail` to 12991 which drops to 229 m^3

notes:

missing two cruise: 
ok - 2016-11 MR 64 - because no record out volume

issue - 2016-03 all stations and mesh have duplicate matches from meta to one micro sheet

ok - 2017-01 WS 64 - has a duplicate samples for some reason
ok - 2015-04 WS 500 - has a duplicate samples for some reason

NOTE: cruise March 2016 still need some work because of the duplicate
 
```{r}
species_unnest %>%
  nest(.by = sheet_nm:meta, .key = "nested_data") %>%
  filter(
    between(row_number(), 52, 52 + 8)
  )


# merge metadata with zoo data
# this combines month, year, station and mesh 
species_merged_meta <-
  species_unnest %>%
  nest(.by = sheet_nm:meta, .key = "nested_data") %>%
  filter(
    !between(row_number(), 52, 52 + 8) # March 2016 - duplicate issue
  ) %T>%
  print() %>%
  # left_join(
  full_join(
    meta_df,
    by = join_by(
      year,
      month,
      "stn" == "locationID",
      "mesh" == "mesh_size_um"
    ),
    # relationship = "one-to-one"
  ) %>%
  clean_names() %T>%
  print()

species_merged_filt <- 
  species_merged_meta %>%
  filter(!is.na(sheet_nm)) %>%
  rename("split_size_og" = split_size) %>%
  unnest(nested_data) %>%
  janitor::remove_empty("cols")

get_dupes(species_merged_filt)
visdat::vis_miss(species_merged_filt)


#  LK071664	2016	7	LK	64	
# filter(meta_df, row_number() == 114)
```



```{r}

species_merged_filt %>%
  distinct(sheet_nm, .keep_all = TRUE) %>%
  select(1:6, "old_vol" = vol_filtered_m3, "new_vol" = volume_filt_cubic_m) %>%
  mutate(
    diff = new_vol - old_vol, 
    matters = if_else(abs(diff) < 1, "no", "yesssssss"),
    diff = format(diff, scientific = FALSE)
  )  %T>% 
  print() %>%
  count(matters)


```


# Save Files


```{r save-files}
species_unnest %>%
  janitor::remove_empty("cols") %>%
  select(-meta) %>%
  save_csv(
    .data          = .,
    save_location  = here(base_dir, "processed"),
    save_name      = "zoo_compiled",
    # overwrite      = TRUE, # comment in when overwriting
    verbose        = TRUE,
    time_stamp_fmt = "%Y%m%d"
)

species_merged_filt %>%
  janitor::remove_empty("cols") %>%
  select(-meta) %>%
  save_csv(
    .data          = .,
    save_location  = here(base_dir, "processed"),
    save_name      = "zoo_compiled_with_meta",
    # overwrite      = TRUE, # comment in when overwriting
    verbose        = TRUE,
    time_stamp_fmt = "%Y%m%d"
)

```

# Read 2015 - 2017 OBIS Files (needs fixing)
```{r}

"https://ipt-obis.gbif.us/archive.do?r=sfmbon_zooplankton&v=1.5"
sf_mbon <- 
    here("data", "example", 
         "claudia_zoo_biometry",
         "dwca-sfmbon_zooplankton-v1.5", "occurrence.txt") %>%
    read_delim(
        delim = "\t", 
        escape_double = FALSE, 
        trim_ws = TRUE,
        show_col_types = FALSE)

sf_mbon %>%
    filter(
        !eventDate > as.Date("2017-01-01")
    ) %>%
      View()

sf_mbon %$% unique(eventDate) %>%
    as.Date() %>%
    sort()

here("data", "example", "claudia_zoo_biometry", 
     "0122465-220831081235567", "verbatim.txt") %>%
    read_delim(.,
               delim = "\t", 
               escape_double = FALSE,
               trim_ws = TRUE) %>%
    remove_empty(which = "cols") %>%
    View()

here("data", "example", "claudia_zoo_biometry", 
     "0122465-220831081235567", "occurrence.txt") %>%
    read_delim(.,
               delim = "\t", 
               escape_double = FALSE,
               trim_ws = TRUE,
               show_col_types = FALSE) %>%
    remove_empty(which = "cols") %>%
    select(1:day) %>%
    View()



```

```{r}
species_merged_filt2 <- 
  here(base_dir, "processed") %>%
  dir_ls(regexp = "zoo_compiled_with_meta") %>%
  last_mod() %>%
  read_csv(show_col_types = FALSE)
```

```{r}
new_data <- 
  here("data", "processed", "pre_obis") %>%
    dir_ls(regexp = "zoo_data_pre_obis") %>%
    .[10] %>%

    # last_mod() %>%
    read_csv(show_col_types = FALSE) %>%
    filter(
      !str_detect(mesh, "64") 
      & !str_detect(site, "9B|57")
    ) 


new_data_summary <- 
  new_data %>%
  summarise(
    .by = c(cruise_id, site, mesh, date_time),
    
    new_ind = sum(number_ind_sample, na.rm = TRUE)
    
  ) %>%
    rename("stn" = site)
new_data_summary2 <- 
  new_data %>%
  summarise(
    .by = c(cruise_id, site, mesh, date_time),
    
    new_ind = sum(ind_m3, na.rm = TRUE)
    
  ) %>%
    rename("stn" = site)

```


```{r}
species_merged_filt2 %>%
  summarise(
    .by = c(sheet_nm, cruise_id, stn, mesh, date_time),
    
    old_ind = sum(total_ind_sample, na.rm = TRUE),
    new_ind = sum(total_ind_correct, na.rm = TRUE)
    
  ) %>%
  ggplot(aes(x = date_time)) +
    geom_line(aes(y = old_ind), color = "red") +
    geom_line(aes(y = new_ind), color = "blue") +
    geom_point(aes(y = old_ind), color = "red") +
    geom_point(aes(y = new_ind), color = "blue") +
    theme_minimal() + 
    facet_wrap(~mesh + stn, ncol = 3, scale = "free_y")
```


```{r}
species_merged_filt2 %>%
  summarise(
    .by = c(sheet_nm, cruise_id, stn, mesh, date_time),
    
    old_ind = sum(total_ind_sample, na.rm = TRUE),
    new_ind = sum(total_ind_correct, na.rm = TRUE)
    
  ) %>%
  ggplot(aes(x = date_time)) +
    geom_line(aes(y = old_ind), color = "red") +
    geom_line(aes(y = new_ind), color = "blue") +
    geom_line(
      data = new_data, aes(y = new_ind), color = "blue") +
    geom_point(aes(y = old_ind), color = "red") +
    geom_point(aes(y = new_ind), color = "blue") +
    theme_minimal() + 
    facet_wrap(~mesh + stn, ncol = 3, scale = "free_y")
```
```{r}
species_merged_filt2 %>%
  mutate(
    vol_filt_new = total_ind_correct / volume_filt_cubic_m
  )  %>%
  summarise(
    .by = c(sheet_nm, cruise_id, stn, mesh, date_time),
    
    new_ind = sum(vol_filt_new, na.rm = TRUE)
    
  ) %>%
  ggplot(aes(x = date_time)) +
    # geom_line(aes(y = old_ind), color = "red") +
    geom_line(aes(y = new_ind), color = "blue") +
    geom_line(
      data = new_data_summary2, aes(y = new_ind), color = "blue") +
    # geom_point(aes(y = old_ind), color = "red") +
    geom_point(aes(y = new_ind), color = "blue") +
    coord_cartesian(ylim = c(-5, NA)) +
    theme_minimal() + 
    facet_wrap(~mesh + stn, ncol = 3, scale = "free_y")
```





