---
title: "Fix and Convert Previous Zooplankton Data"
author: "Tylar Murray"
date: "2023-04-14"
format: html
---

# 1.0 ---- Setup ----
## 1.1 Load Libraries
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
if (!nzchar(system.file(package = "librarian"))) 
    install.packages("librarian")

librarian::shelf(
    librarian, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    forcats, lubridate, glue, fs, magrittr, here, conflicted,
    # broom # optional
    
    # additional
    cli, naniar, Hmisc
)

# shelf(conflicted) # may be needed if won't allow loading of certain packages

conflicts_prefer(
    dplyr::filter(), 
    dplyr::select()
    )
```

# 1.1 Make sure location of files is correct
From fix_prev_zoo_data.qmd
- create folders
  - place original data
  - place processed file to load here 
  
In case, you don't have these folder, creates them here
```{r create-folder}
base_dir <- here::here("data", "zoo_pre_2018")

fs::dir_create(here::here(base_dir, "processed"))
fs::dir_create(here::here(base_dir, "orig"))

```

## 1.2 Get File Paths
sp.csv from:
<https://drive.google.com/drive/folders/1IKBqpgOR6-bhJnm3wHwZAzvSYAA24Bsx>
put in "~/data/zoo_pre_2018"

```{r get-data-path}
# get file path to processed data
file_path <-
  here::here(base_dir, "processed") %>%
  fs::dir_ls(
    regexp  = "zoo_compiled_with_meta",
    recurse = TRUE
  ) %>%
  # function to check more recent made file to use
  last_mod() 

# read `.csv` file created by `fix_prev_zoo_data.Rmd`
dat <-
  file_path %>%
  # "../zoo_compiled_with_meta_20230411.csv" %>%
  
    readr::read_csv(
    file        = .,
    show_col_types = FALSE,
    guess_max   = 10000,
    name_repair = janitor::make_clean_names
  )

species_path <-
    here::here(base_dir) %>%
    fs::dir_ls(regexp = "sp\\.csv")

# read manually created verbatim name map `sp.csv`
species_map <-
  species_path %>%
  # "../sp.csv" %>%
  
  readr::read_csv(
    file = .,
    show_col_types = FALSE,
    name_repair    = janitor::make_clean_names,
    na = c("", "NA", stringi::stri_dup("?", 1:5)) # ignore `?` 1 to 5 times
  )
```

### 1.2.2 Add Column if Needed

I don't remember how you modified `sp.csv` so I add the column if doesn't exist
```{r}
species_map <-
  species_map %>%
  {
    if (!"jaimes_taxa_info" %in% names(.)) {
      mutate(.,
        jaimes_taxa_info = classification,
        .after = classification
      )
    } else {
      .
    }
  }
```

# 2.0 ---- Merge Taxa Info ----
## 2.1 Merge Species Info with Raw Data
```{r load-csv-data}
# map Jaime's species names to WoRMS-readable using `sp.csv`
dat_merg <-
  # dplyr::select(species_map, 1:2, 4) %>% # select 1st 2 columns
  dplyr::select(species_map,
                classification,
                jaimes_taxa_info,
                sci_name,
                lifestage) %>% # select 1st 2 columns

  # join on classification columns. `.` is where it gets piped into
  dplyr::left_join(
    dat,
    .,
    by = c("clasification" = "jaimes_taxa_info")
  ) %>%
  # construct unique identifier for each row (used for occurrenceID later)
  dplyr::mutate(
    orig_data_row_ID = glue("{sheet_nm}:{clasification}:{lifestage}"),
  ) %>%
    distinct()
```

## 2.2 Check Duplicates for Unique IDs
```{r check-duplicates}
# validate: rows in file must be unique
# janitor::get_dupes(<dataframe name>, <column names to group together>) 

dupes <- janitor::get_dupes(dat_merg, orig_data_row_ID)
# dupes <- 
    janitor::get_dupes(dat, sheet_nm, clasification)
dupes
unique(dupes$sheet_nm)
unique(dat_merg$clasification)
# TODO: how to throw an error and not continue running if dupes exist?
#       pseudocode:
# if(len(dupes) > 0){
#     raise AssertionError(f"duplicate rows found:\n\t {dupes}")
# }

# drop duplicate rows

dat2 <- unique(dat_merg)
dupes2 <- janitor::get_dupes(dat2, orig_data_row_ID)

# TODO: some duplicates still exist. 
# only flowmeter data differs and one looks unrealistically large.

dat2 <- 
    dat2 %>%
    {
    if (!"sci_name_guessed_by_tylar" %in% names(.)) {
      mutate(.,
        sci_name_guessed_by_tylar = classification,
        .after = classification
      )
    } else {
      .
    }
  }

```

# 3.0 ---- Get AphiaID from WoRMS ----
Using `merge_taxa()` 
check = if want to check unmatched names
viewer = if want to see unmatched taxa using View()
```{r aphia-id}
# set location for aphiaID list and the base of the file name
file_exprs <- 
    file_expr(
    # loc       = here::here("data", "metadata", "aphia_id"), 
    loc       = here::here(base_dir), 
    file_base = "aphia_taxa_jaime"
    )

taxa <-
    dat2 %>%
    select("sci_name_guessed_by_tylar") %>%
    distinct() %>%
    merge_taxa(., 
               .file_expr = file_exprs, 
               check      = FALSE, 
               use_cloud = FALSE
               # viewer     = TRUE, 
               )

dat2 <- 
    # select(taxa, 1, 4, 5)  %>%  # select 1st 2 columns
    select(taxa, 1, 4, 5)  %>%  # select 1st 2 columns
  right_join(
      dat2,
      .,
      by = c("sci_name_guessed_by_tylar" = "taxa_orig")
  )  # join on classification columns. `.` is where it gets piped into

# round individual count. we don't know why there are fractional counts. 
dat2 <- mutate(
    dat2,
    ind_count = round(ind_count),
    # hardcode missing lat lon using averages in water-samples
    lon_in = case_when(
        is.na(lon_in) & str_detect(stn, "LK") ~ -81.41464542,
        is.na(lon_in) & str_detect(stn, "WS") ~ -81.71532752,
        is.na(lon_in) & str_detect(stn, "MR") ~ -80.37953386,
        TRUE ~ lon_in
    ),
    lat_in = case_when(
        is.na(lat_in) & str_detect(stn, "LK") ~ 24.53862793,
        is.na(lat_in) & str_detect(stn, "WS") ~ 24.47605506,
        is.na(lat_in) & str_detect(stn, "MR") ~ 25.00607727,
        TRUE ~ lat_in
    )
)

# === set up to align with next chunk var names
dat2 <-
  rename(
    dat2,
    "taxa_orig"        = sci_name_guessed_by_tylar,
    "individualCount"  = ind_count,
    "lifeStage"        = lifestage,
    "decimalLatitude"  = lat_in,
    "decimalLongitude" = lon_in,
  )

eventID   <- "jaime_protocol_observation"
datasetID <- "USF_IMaRS_MBON_compiled_zoo_taxonomy_jaimie_2018"

```


```{r construct occurrence table}
names(dat2)
# map col names into dwc
occur <-
    dat2 %>%
    # left_join(event, 
    #           by = c("cruise_id", 
    #                  "site" = "locationID", 
    #                  "mesh" #, 
    #                  # "recordedBy", 
    #                  #"recordedByID", "basisOfRecord"
    #                  )
    #           ) %>%
    dplyr::transmute(
    # this method is prefered to maintainers instead of transmute
    # dplyr::mutate(
        # .keep = "used",
        # ---
        
        decimalLatitude,
        decimalLongitude,
        # eventID, 
        # eventDate                = glue("{date}T{local_time_est}ET"),
        # make datetime and convert to utc 
        eventDate = lubridate::ymd_hms(stringr::str_c(date, local_time_est),  
                                       tz = "EST"),
        eventDate = lubridate::format_ISO8601(eventDate, usetz = "Z"),
        
        
        # occurrenceID             = glue("{eventID}:{aphiaID}:{lifeStage}"),
        # changed to taxa_orig to make sure occur ID is unique
        occurrenceID             = glue("{eventDate}:{eventID}:{taxa_orig}:{lifeStage}"),
        occurrenceID = str_replace_all(occurrenceID, " ", "_"),
        # taxa names
        scientificName,
        # across(kingdom:genus),
        # taxonRank = rank,
        # taxonID = taxon_rank_id,
         
        # recorded by
        # recordedBy,
        # recordedByID,
        # dateIdentified           = date_analyzed,
        
        # counts of organisms?
        # IDk which of these is accurate
        # individualCount          = number_ind_sample,
        # add each aliquot as a total count per sample event
        individualCount,
        # organismQuantity         = ind_m3,
        # organismQuantityType     = "Individuals per cubic metre",
        # or
        # measurementType          = "Number per cubic metre",
        # MOF? should also include ind/sample?
        # measurementUnitID        = 
        # "http://vocab.nerc.ac.uk/collection/P06/current/UPMM/",
        # MOF?
        
        # TODO: check this information ----
        lifeStage,
        # establishmentMeans       = "native | uncertain",
        occurrenceStatus         = "present",
        preparations             = "formalin before analysis | ethanol after analysis",
        scientificNameID,
        # like urn:lsid:ipni.org:names:37829-1:1.3
        basisOfRecord = "humanObservation",
        datasetID = {datasetID},
        identificationReferences = "WoRMS",
        verbatimIdentification   = taxa_orig,
        # georeferenceVerificationStatusProperty = "verified by contributor",
        georeferenceVerificationStatus = "verified by contributor",
        dispostion = "in collection",
        coordinateUncertaintyInMeters = 5,
        minimumDepthInMeters = 0,
        maximumDepthInMeters = 2,
    )

occur <- filter(occur, !is.na(scientificNameID))
naniar::vis_miss(occur)
Hmisc::describe(occur)


if (!file.exists(here(base_dir, "occurrences.csv")))
    readr::write_csv(occur, here::here(base_dir, "processed", "occurrences.csv"))
```