---
title: "Setup Project"
author: "Sebastian DiGeronimo"
date: "2022-11-22"
output: html_document
---

# Step 1: Load/install most used packages 
```{r setup, include=FALSE}
library("ggplot2")
library("tibble")
library("tidyr")
library("readr")
library("purrr")
library("dplyr")
library("stringr")
library("forcats")
library("lubridate")
library("glue")
library("fs")
library("magrittr")
library("here")
```

# Step 2: Source scripts with functions needed and attach to search path

1. Creates directory for consistent file structures

2. Copy files from designated location
    - The first time using, this will open a `.Rprofile` file. This will be used to create a variable called `box_dir`. This is where the raw abundance directory is. If you want to select through popup set `copy_files_box(.choose = TRUE)` and will paste to console. You can then copy and paste it into the .Rprofile. You will then need to re-source the chunk.
    - If you want to auto-look for new files, copy and paste the chunck `funcs` into the `.Rprofile` and each `Restart` in R will look for new files and ask if you want to download. 
        - You can either:
            1. `ask` = `TRUE` for asking if you want to copy and which ones
            2. `auto` = `TRUE` to copy any file that is new

3. Load functions for searching the raw for taxa names and connecting it to a master sheet
- if a master sheet doesn't exists locally, it will start a new one and put it in:
    - `data/metadata/aphia_id`

```{r funcs}
# source scripts with functions in new environment
my_funcs <- new.env()

# create directory structure if does not exists and create tree map
source(here::here("scripts","create_dir.R"), local = my_funcs)
# copy files from box
source(here::here("scripts", "copy_files_from_box.R"), local = my_funcs)
# loading check for taxa
source(here::here("scripts", "taxa_list_check.R"), local = my_funcs)


attach(my_funcs)
rm(my_funcs)

copy_files_box(ask = TRUE)

```
