---
title: "Taxonomix Matchup to WoRMs"
author: "Sebastian DiGeronimo"
date: "3/2/2022"
output: html_document
---
# Load Packages
This loads any packages that are not contained in the .Rprofile.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
librarian::shelf(
    # librarian, ggplot2, tibble, tidyr, readr, purrr, dplyr, stringr,
    # forcats, lubridate, glue, fs, magrittr, here,
    # broom, # optional
    
    readxl, hablar, worrms, ggforce, geosphere, vroom, plotly,
    
    quiet = TRUE
    )
```

# Load functions if not already
This will load the functions needed if not already contain in the search path.
```{r load-functions}
if (!rlang::is_attached('my_funcs')) { 
    cli_alert_info(
        c("Attaching project functions to the {.code searchpath} as ",
          "{.code my_funcs}."))
    
    # source scripts with functions in new environment
    my_funcs <- new.env()
    
    # create directory structure if does not exists and create tree map
    source(here::here("scripts","create_dir.R"), local = my_funcs)
    # copy files from a cloud service
    source(here::here("scripts", "copy_files_from_cloud.R"), local = my_funcs)
    # loading check for taxa 
    source(here::here("scripts", "taxa_list_check.R"), local = my_funcs)
    
    attach(my_funcs)
    rm(my_funcs)
}
```

# Load Data Sets
1. Parse all file paths in `data/raw/`. 
- Will ignore files with `~$` in front becasue these are usually temporary files that are opened
2. `skip_file` will check a processed file to ignore files that were previously loaded and formatted
3. Take leftover files and merge data into long format
- files contain some metadata for each sample processed and taxonomic data
```{r load-taxa-data}
# ---- find all file names and filter ----
file.taxa <- 
    fs::dir_ls(path =  here::here("data", "raw"),
               regexp = "^[^~]*\\.xlsx$",
               recurse = TRUE) %>%
    skip_file(.)

# ---- load all files into one tibble ----
taxa_data <-
    file.taxa %>%
    map_dfr(., ~ load_data(.x), .id = "Files") %>%
    mutate(Files = basename(Files))
taxa_data

```
# Merge taxanomic list and save new data
1. Loads previous master taxonomic data sheet
- This will check if it is contained in the cloud service at the root level, or create one if it doesn't exists
- If it doesn't exist, it will check the long format data for a column called taxa
- Then, it will `clean` the taxa list by removing sp., spp. and life stage information
- Next, it will use the custom function for searching scientific names in `WoRMS`
    - The base for this is `obistools::match_taxa`, but it adds the ability to type non-matched names or to use common names
- Once completed, it will save the taxa list in `data/metadata/aphia_id/` 
- It will check if there are any `NAs` if set `check = TRUE`
    - Only the `NA` rows will be rechecked again. Then it will merge with previous sheet with a new filename 

2. Merge taxa lists with long format abundance data.
- Taxa list contains `taxa_orig`, `taxa`, `lifeStage`, `scientificName`, `scientificNameID`, `match_type`

3. Save merged data in merged long format and separate samples files in `data/processed/` and `data/processed/ind_file_merg/` 
- If supplied `file.taxa`, it will append `processed_files.csv` (the list of files) to skip in future runs in `data/metadata/`
```{r}
# TODO: read in all files, joining with scientific names, then add more if they 
# don't exist
taxa_matched_merg <- merge_taxa(taxa_data, check = TRUE) %>%
    relocate(Files, cruise_id, date_collected, site, .before = 1) %>%
    arrange(date_collected, site, taxa)

save_merged(taxa_matched_merg = taxa_matched_merg, 
            .taxa_file = file.taxa
            )

```

# Push current master taxonomic sheet to cloud
1. Load most recent taxa sheet
2. Check if there any unmatched taxa
3. Push to cloud location
```{r push-to-cloud}
taxa_list <- load_taxa_list(check = TRUE) 
no_matches <- taxa_list %>%
    filter(is.na(scientificName))


master_taxa_cloud(taxa_list  = taxa_list, 
                  .cloud_dir = cloud_dir, 
                  where_to   = "cloud")
```

# tried a different database without working
```{r}
# library("taxize")
# 
# 
# 
# cbind(taxa.name, test)
# 
# ?get_tsn
# get_tsn(taxa.name[1:5])
# get_tsn(("Acartia"))
# get_wormsid(test)
# get_wormsid("fish", searchtype = 'common')
# 
# rgbif::name_backbone("Acartia")
# 
# obistools::match_taxa(c("Acartia","Candacia","copepod","Euterpina acuntifrons"), ask = T)
```

```{r}


# obis.taxa <- obistools::match_taxa(taxa.name, ask = T)
# 
# robis::taxon(11676)
# 
# # with this 20 have NA without ability to ask for input
# obis.taxa %>%
#     mutate(taxa.name, .before = 1)
```

